\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.4cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{booktabs,enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,fit,backgrounds}
\definecolor{tierspend}{RGB}{200,60,60}
\definecolor{tierfull}{RGB}{210,140,40}
\definecolor{tierin}{RGB}{60,130,180}
\definecolor{tierrecv}{RGB}{70,150,90}
\usepackage[colorlinks=true,linkcolor=blue!50!black,citecolor=blue!50!black,urlcolor=blue!50!black]{hyperref}

\emergencystretch=3em
\hbadness=10000
\hfuzz=2pt

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{construction}[theorem]{Construction}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\newcommand{\F}{\mathbb{F}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\renewcommand{\Pr}{\mathrm{Pr}}
\newcommand{\Fp}{\F_p}
\newcommand{\Fq}{\F_q}
\newcommand{\ip}[2]{\langle #1,\, #2 \rangle}
\newcommand{\Comm}{\mathsf{Commit}}
\newcommand{\Open}{\mathsf{Open}}
\newcommand{\Setup}{\mathsf{Setup}}
\newcommand{\Verify}{\mathsf{Verify}}
\newcommand{\negl}{\mathsf{negl}}
\newcommand{\poly}{\mathsf{poly}}
\newcommand{\PPT}{\mathsf{PPT}}
\newcommand{\Adv}{\mathcal{A}}
\newcommand{\Prov}{\mathcal{P}}
\newcommand{\Ver}{\mathcal{V}}
\newcommand{\Sim}{\mathcal{S}}
\newcommand{\Ext}{\mathcal{E}}
\newcommand{\Rel}{\mathcal{R}}
\newcommand{\Lang}{\mathcal{L}}
\newcommand{\nfsym}{\mathsf{nf}}
\newcommand{\cmsym}{\mathsf{cm}}
\newcommand{\cmx}{\mathsf{cmx}}
\newcommand{\cvsym}{\mathsf{cv}}
\newcommand{\rk}{\mathsf{rk}}
\newcommand{\Pallas}{\ensuremath{\mathsf{Pallas}}}
\newcommand{\Vesta}{\ensuremath{\mathsf{Vesta}}}
\newcommand{\Ep}{\ensuremath{\mathcal{E}}}
\newcommand{\NoteCommit}{\ensuremath{\mathsf{NoteCommit}}}
\newcommand{\Sinsemilla}{\ensuremath{\mathsf{Sinsemilla}}}
\newcommand{\ShortCommit}{\ensuremath{\mathsf{ShortCommit}}}
\newcommand{\Extract}{\ensuremath{\mathsf{Extract}}}
\newcommand{\Acc}{\ensuremath{\mathsf{Acc}}}
\newcommand{\GroupHash}{\ensuremath{\mathsf{GroupHash}}}
\newcommand{\ask}{\ensuremath{\mathsf{ask}}}
\newcommand{\aksym}{\ensuremath{\mathsf{ak}}}
\newcommand{\nksym}{\ensuremath{\mathsf{nk}}}
\newcommand{\fvk}{\ensuremath{\mathsf{fvk}}}
\newcommand{\ovk}{\ensuremath{\mathsf{ovk}}}
\newcommand{\dk}{\ensuremath{\mathsf{dk}}}
\newcommand{\pkd}{\ensuremath{\mathsf{pk_d}}}
\newcommand{\gd}{\ensuremath{\mathsf{g_d}}}
\newcommand{\rcv}{\ensuremath{\mathsf{rcv}}}
\newcommand{\rcm}{\ensuremath{\mathsf{rcm}}}
\newcommand{\rsk}{\ensuremath{\mathsf{rsk}}}

\renewenvironment{abstract}{%
  \small\begin{center}{\bfseries\abstractname}\end{center}\medskip\noindent\ignorespaces
}{\par\medskip}

\title{\textbf{\Huge Halo 2 Guide}\\[6pt]\large The Halo 2 proof system, from arithmetisation to recursion}
\author{m@rek.onl}
\date{}
\begin{document}
\maketitle
\thispagestyle{empty}
\begin{abstract}
This is the third volume of a four-part series --- \emph{Math Guide}, \emph{Crypto Guide}, \emph{Halo~2 Guide}, \emph{Orchard Guide}. Assuming the mathematics of the \emph{Math Guide} and the cryptographic primitives of the \emph{Crypto Guide}, it develops the Halo~2 proof system: the polynomial-IOP design pattern, a rigorous account of knowledge soundness and the algebraic group model, recursive proof composition and accumulation schemes, and finally Halo~2 itself --- PLONKish arithmetisation, the permutation and lookup arguments, the inner-product polynomial commitment and the Halo accumulation trick, and the complete protocol.
\end{abstract}
\clearpage
\tableofcontents
\clearpage

\section{Introduction}\label{sec:intro}
This is the \emph{Halo~2 Guide}, the third volume of a four-part development of the cryptography
behind Zcash's Orchard protocol. The first two volumes --- the \emph{Math Guide} and the
\emph{Crypto Guide} --- constructed, from the ground up, the mathematics (finite fields, elliptic curves,
polynomials and evaluation domains, probability) and the cryptographic primitives (hash functions,
commitments, $\Sigma$-protocols and the Fiat--Shamir transform, zero knowledge, polynomial
commitment schemes) that this volume assumes. The present volume assembles those ingredients into a
\emph{proof system}.

The development proceeds in four steps. First, we make the polynomial-IOP design pattern precise: the
recipe ``encode a computation as polynomial identities over an evaluation domain, then compile to a
succinct argument with a polynomial commitment.'' Next, we treat knowledge soundness rigorously,
including the algebraic group model and the tree-of-transcripts extractor that the inner-product
argument requires. We then motivate recursive proof composition and develop the accumulation
schemes that make it practical. Finally, we assemble Halo~2 itself:
PLONKish arithmetisation, the permutation and lookup arguments, the inner-product polynomial
commitment with a worked example and its accumulation (the ``Halo trick''), and the complete
seven-phase protocol. The \emph{Orchard Guide} subsequently uses this proof system to construct a shielded
payment protocol.

\section{Polynomial IOPs and the SNARK design pattern}\label{sec:piop}

The modern SNARK design pattern factors the protocol into two clean
layers: a \emph{polynomial interactive oracle proof} (PIOP) for the
relation, and a polynomial commitment scheme (the \emph{Crypto Guide}) that
realises the PIOP's oracles. This factoring, due in its current form to
Bünz, Fisch, Szepieniec, and others, has substantially clarified the
nature of SNARKs.

\subsection{The polynomial IOP, formally}

\begin{definition}[Polynomial IOP]\label{def:piop}
A \emph{polynomial interactive oracle proof} for a relation $\Rel$ is a
public-coin interactive protocol where in each round:
\begin{itemize}[leftmargin=*]
\item The prover sends one or more \emph{polynomial oracles} $p_i \in \F[X]$
(each of bounded degree).
\item The verifier sends a random challenge $c \in \F$.
\end{itemize}
At the end of the protocol, the verifier may query each oracle at a small
number of points, learning the polynomial's evaluations at those points.
The verifier's final accept/reject is a deterministic function of the
challenges, evaluations, and the public input $x$.

We define completeness, soundness, and knowledge soundness as usual.
\end{definition}

A PIOP differs from a vanilla interactive argument in one essential respect:
the verifier does not read the prover's messages in full --- it queries
only a few evaluations of each polynomial oracle. This is what renders the
verifier ``succinct'' even when the polynomials themselves have large degree.

\subsection{Compiling a PIOP into a SNARK}

\begin{construction}[PIOP-to-SNARK compiler]\label{const:piop-snark}
Given a PIOP $\Pi$ for $\Rel$, a PCS scheme $(\Setup, \Comm, \Open, \Verify)$
for $\F[X]$, and a random oracle $H$, the compiler produces the SNARK:
\begin{enumerate}[leftmargin=*]
\item Setup: run the PCS setup to obtain $\mathsf{pp}$.
\item Prover: simulate $\Pi$, replacing each polynomial oracle $p_i$ by its
PCS commitment $C_i$, and each verifier-challenge / oracle-query round by
the corresponding PCS opening. Use Fiat--Shamir (the \emph{Crypto Guide}) to derive
challenges from the transcript.
\item Verifier: parse the transcript, recompute Fiat--Shamir challenges,
verify each PCS opening, and run the PIOP's final accept/reject check.
\end{enumerate}
\end{construction}

\begin{theorem}[Soundness preservation]\label{thm:piop-snark}
If $\Pi$ has round-by-round (equivalently, state-restoration) knowledge
soundness $\kappa_{\Pi}$ and the PCS is extractable
with extraction error $\kappa_{\mathrm{PCS}}$ (the PCS definition of the
\emph{Crypto Guide}), the compiled SNARK has knowledge soundness
$O(Q \cdot (\kappa_\Pi + \kappa_{\mathrm{PCS}}))$ in the random oracle model,
where $Q$ is the adversary's RO-query budget.
\end{theorem}

The query budget multiplies $\kappa_\Pi$ as well as
$\kappa_{\mathrm{PCS}}$: step 2 derives the PIOP challenges by
Fiat--Shamir, so a cheating prover that re-randomises a commitment
redraws every subsequent challenge, obtaining up to $Q$ independent
draws of each --- the grinding attack. Round-by-round soundness confines
this loss to the single factor $Q$; for an $r$-round PIOP with plain
knowledge soundness only, the naive Fiat--Shamir analysis loses a factor
$Q^{O(r)}$ (\S\ref{sec:ks}).

This factoring is of considerable utility: research on PIOPs (PLONK, Marlin,
Spartan, HyperPlonk, $\dots$) proceeds independently of research on
polynomial commitment schemes, and any compatible pair combines into a new
SNARK that inherits the PIOP's relation and the commitment scheme's
cryptographic profile. This document
develops the one PCS that Halo 2 uses --- the inner-product argument
(\S\ref{sec:ipa-detail}) --- and treats the PIOP/PCS split only insofar as
it clarifies the Halo 2 construction.

\subsection{A worked example: a PIOP for univariate evaluation}\label{ex:univariate-piop}

To render the abstraction concrete, consider a minimal PIOP for the simplest
non-trivial relation: ``the prover knows a polynomial $p$ of degree $\le d$
such that $p(z) = y$ for a public $(z, y)$.''

\begin{construction}[PIOP for univariate evaluation]\label{const:univ-piop}
On input $(z, y)$ and prover witness $p$:
\begin{enumerate}[leftmargin=*]
\item Prover sends the polynomial oracle $p$ and the quotient oracle
$q := (p - y)/(X - z)$.
\item Verifier samples a random challenge $r \in \F$.
\item Verifier queries $p$ and $q$ at $r$.
\item Verifier accepts iff $p(r) - y = q(r) \cdot (r - z)$.
\end{enumerate}
\end{construction}

Completeness is immediate ($q$ is a polynomial since $p(z) - y = 0$).
Soundness follows from Schwartz--Zippel: if $(X - z) \nmid (p(X) - y)$,
then $p(X) - y - q(X)(X - z)$ is a non-zero polynomial of degree $\le d$,
and its random evaluation is zero with probability $\le d/|\F|$.

Compiling this PIOP with the IPA-PCS, applying Fiat--Shamir, yields a
non-interactive SNARK for the same relation. The PCS handles polynomial
hiding and binding; the PIOP handles the relation logic. This separation
is precisely the modular SNARK design pattern.

\subsection{PLONK as a PIOP}

PLONK (\S\ref{sec:plonkish}) is a more elaborate PIOP. Its
oracles are:
\begin{itemize}[leftmargin=*]
\item One advice polynomial per advice column.
\item A permutation running-product polynomial $Z_{\mathrm{perm}}$.
\item Two permuted lookup columns \(A'\), \(S'\) per lookup.
\item One lookup running-product polynomial per lookup.
\item A quotient polynomial $h$ (split into chunks).
\end{itemize}
Its challenges are, in order: $\theta$ (lookup column compression),
$\beta, \gamma$ (the permutation and lookup products), $y$ (the
gate-equation combination), and $x$ (the final evaluation point). Its final check is the
PLONK gate equation $C(x) = h(x) \cdot Z_H(x)$, which the verifier checks from
the evaluations of the oracles at $x$ (and possibly $\omega x$).

This is precisely the PIOP that phases 1--5 of the Halo 2
protocol of \S\ref{sec:halo2-full} compile, phase 5's claimed evaluations
answering its oracle queries; phases 6--7 bind those answers to the
commitments --- the multipoint reduction (the challenges
$x_1, \dots, x_4$ and the combined polynomial $f^*$) followed by a single
IPA-PCS opening of $f^*$.

\section{A rigorous treatment of knowledge soundness}\label{sec:ks}

We now develop knowledge soundness with the care needed for arguments
about modern SNARKs. The treatment combines the rewinding-extractor
framework (the \emph{Crypto Guide}) with the algebraic group model.

\subsection{Definitional refinements}

The bare definition of the \emph{Crypto Guide} suffices for $\Sigma$-protocols
but is awkward for compound systems. We record the following standard refinement.

\begin{remark}[Witness-extended emulation]\label{def:wee}
An interactive argument $(\Prov, \Ver)$ has \emph{witness-extended
emulation} if there is a $\PPT$ emulator $\Ext$ such that for every $\PPT$
prover $\Prov^*$ and every input $x$:
\begin{itemize}[leftmargin=*]
\item Run $\Prov^*$ honestly with $\Ver$; let $\tau$ be the transcript and
$b$ the verifier's accept/reject bit.
\item Run $\Ext^{\Prov^*}(x)$; it outputs $(\tau', b', w)$.
\end{itemize}
The distributions of $(\tau, b)$ and $(\tau', b')$ are computationally
indistinguishable, and moreover whenever $b' = 1$, $(x, w) \in \Rel$ except
with negligible probability.
\end{remark}

WEE captures simultaneously two properties: ``the extractor produces a witness whenever
the protocol accepts'' and ``the extractor's external behaviour resembles
a normal protocol execution.'' It is the appropriate notion for arguments
embedded in larger protocols, such as the batched verification of many
proofs at once (\S\ref{sec:halo-acc}); plain knowledge soundness,
however, suffices for this volume's statements.

\subsection{The algebraic group model}

The IPA's knowledge-soundness proof proceeds through the tree extractor of
\S\ref{sec:ipa-detail}, with knowledge error $O(d / |\F|)$. The
interactive tree extractor costs $3^{\log_2 d} = d^{\log_2 3}$
transcripts and runs in expected polynomial time
(\S\ref{sec:tree-extract}); the looseness arises for the non-interactive
argument, where the naive Fiat--Shamir analysis of the $\log d$-round
protocol loses a factor $q^{O(\log d)}$ in the adversary's random-oracle
query count $q$ --- quasi-polynomial for polynomial $q$.
The \emph{algebraic group model} (AGM; Fuchsbauer, Kiltz, Loss; CRYPTO
2018) removes this loss by restricting the adversary's group operations.

\begin{definition}[Algebraic adversary]\label{def:algebraic}
An \emph{algebraic adversary} against a group $\G$ is a $\PPT$ algorithm
that, whenever it outputs a group element $P \in \G$, additionally outputs
an explicit linear combination
$P = \sum_i [a_i] Q_i$ expressing $P$ as a combination of previously-seen
group elements $Q_i$.
\end{definition}

The AGM sits between the standard model (no restriction on the adversary)
and the generic group model (group operations available only through an
oracle). Most natural adversaries are algebraic --- they compute group
elements by known scalar multiplications and additions --- so the
heuristic is regarded as conservative. In the AGM the forking loss
disappears: an algebraic prover outputs every group element together with
an explicit linear combination of the elements it has seen, so the
extractor reads candidate witnesses directly off these representations
instead of reassembling them from forked transcripts. Ghoshal and Tessaro
(CRYPTO 2021) carry out this analysis for Bulletproofs-style
inner-product arguments, proving tight state-restoration soundness ---
and hence tight Fiat--Shamir security --- in the AGM. The accumulation
analysis of BCMS20 (B\"unz, Chiesa, Mishra, Spooner; TCC 2020), by
contrast, does not invoke the AGM: it proves
security in the random oracle model under the DL assumption
(Theorem~\ref{thm:bcms20-acc}).

\subsection{Tree-extractor analysis}\label{sec:tree-extract}

To render the tree extractor concrete, consider the key step at each round of the
IPA halving (see \S\ref{sec:ipa-detail}):

\begin{lemma}[Single-round extraction]\label{lem:single-round}
In a single round of the IPA at depth $j$, given three accepting
sub-transcripts that share the round's first message $(L_j, R_j)$ but
have three challenges $u_j$ with pairwise distinct squares
($u_j \ne \pm u_j'$), the extractor recovers the round-$j$
witness halves $\vec{a}_{\mathrm{lo}}^{(j)}, \vec{a}_{\mathrm{hi}}^{(j)}$
together with the consistency of the $L_j, R_j$ cross-terms via $3 \times 3$
Vandermonde inversion (rows $(u_j^{-2}, 1, u_j^2)$), since the folded
relation is quadratic in $u_j$.
\end{lemma}

\begin{proof}[Sketch]
The three accepting transcripts give three accepting verifier equations.
Because the folded relation is quadratic in $u_j$ (the three monomials
$u_j^{-2}, 1, u_j^2$), treating these as three linear equations in three
unknowns (the cross-terms and the folded commitment), the system has a
unique solution because the $3 \times 3$ Vandermonde determinant with rows
$(u_j^{-2}, 1, u_j^2)$ is non-zero exactly when the squares $u_j^2$ are
pairwise distinct ($u_j \ne \pm u_j'$). Excluding the bad challenge collisions contributes an additive
$O(1/|\F|)$ term per round to the knowledge error.
\end{proof}

The full extraction requires three accepting transcripts at each of the
$k = \log_2 d$ rounds, giving a tree of $3^k = d^{\log_2 3} \approx d^{1.585}$
leaves --- polynomial in $d$, though more than the $O(d)$ a binary tree
would cost. In the AGM, each prover invocation is direct (the extractor
reads the algebraic representation); in the standard model with forking,
each invocation costs $1/\varepsilon$ expected queries to the prover,
where $\varepsilon$ is the cheating prover's acceptance probability,
yielding total cost $O(d^{\log_2 3} \cdot \mathsf{poly}(\lambda) / \varepsilon)$.

\subsection{Concrete versus asymptotic security}

The literature on knowledge soundness contains numerous tightness gaps. The
more important ones include:
\begin{itemize}[leftmargin=*]
\item Bulletproofs' naive Fiat--Shamir knowledge-soundness analysis
loses a factor $q^{O(\log d)}$, exponential in the round count $\log d$
with base the query count $q$; the interactive standard-model extraction
itself is polynomial ($d^{\log_2 3}$ transcripts, expected polynomial
time).
\item Halo 2's deployed parameters assume the AGM and/or the random oracle
model heuristically.
\end{itemize}

This reflects the practical reality of deployed SNARKs: tight reductions
in idealised models, loose reductions in standard models, and deployment
selecting the former.

\section{Recursive proof composition and accumulation schemes}\label{sec:pcd}

We now turn to recursive proof composition: proving, inside one proof,
that another proof verifies. This is what makes Halo 2's combination of
``no trusted setup'' with recursion possible. Orchard's mainnet
deployment does not recurse, but the accumulation structure shapes the
deployed verifier nonetheless --- the deferred MSM and the batched
verification of \S\ref{sec:halo-acc} --- so we develop it with care.

\subsection{Recursion and its two classical obstacles}

Consider a computation iterated many times,
$z_n = f^{(n)}(z_0) := f(f(\dots f(z_0) \dots))$ --- Valiant's
\emph{incrementally-verifiable computation} (TCC 2008) --- or, more
generally, a DAG of computations each consuming the proofs of its
predecessors --- \emph{proof-carrying data} (Chiesa, Tromer; ICS 2010).
The natural construction is \emph{recursive SNARK composition}: at step
$i + 1$ the prover proves ``$z_{i+1} = f(z_i)$, \emph{and} there exists
a proof $\pi_i$ that the verifier circuit for step $i$ accepts.'' One
proof then attests to the entire history, and the per-step proving cost
is independent of the history's length.

For this to work, the SNARK's verifier must itself be efficiently
expressible as a circuit, and here the classical instantiations meet two
obstacles:
\begin{enumerate}[leftmargin=*]
\item Pairing-based SNARKs have constant-size, constant-time verifiers,
but the verifier evaluates a pairing --- costly to express algebraically
inside a circuit, since it requires extension-field arithmetic --- and
these SNARKs need a trusted setup besides.
\item Transparent SNARKs built on the IPA need no pairing and no trusted
setup, but their verifiers run in linear time --- the $O(d)$ MSM of
\S\ref{sec:gpoly}. A linear-time check dominates the verifier circuit
and negates the per-step efficiency.
\end{enumerate}
Halo (Bowe--Grigg--Hopwood 2019) defeats obstacle 2 directly ---
deferring the linear-time part of verification instead of recomputing it
at every step --- and sidesteps obstacle 1 by using the IPA, which
involves no pairing. BCMS20 formalised the deferral as an
\emph{accumulation scheme}.

\subsection{Accumulation schemes}

\begin{definition}[Accumulation scheme, informal]\label{def:acc}
For a predicate $\Phi$ (for example, ``this IPA opening is correct''), an
\emph{accumulation scheme} consists of:
\begin{itemize}[leftmargin=*]
\item An \emph{accumulator}, a compact data structure representing a batch
of claims about $\Phi$.
\item A \emph{folding} algorithm: given an old accumulator and a new claim
about $\Phi$, it produces a new accumulator that represents both --- the
new accumulator is valid only if the old one was valid and the new claim
holds. Folding is fast: $O(\log d)$ in the IPA instantiation below.
\item A \emph{decider}, which is expensive (linear-time) but which the scheme
invokes only once at
the very end to validate the accumulator. If the decider accepts, every
claim folded in so far is true.
\end{itemize}
\end{definition}

For the IPA-PCS such a scheme exists, and it resolves the recursion
problem in the IPA setting: the in-circuit verifier checks only the
$O(\log d)$ ``succinct'' part of an IPA proof and folds the $O(d)$
``deferred'' MSM check into a running accumulator. The decider pays the
deferred work once, at the end of the entire chain. We state the scheme
precisely as Theorem~\ref{thm:bcms20-acc}, once the IPA construction of
\S\ref{sec:ipa-detail} is in hand.

\subsection{The necessity of a cycle of curves}\label{sec:cycle}

The accumulation scheme's in-circuit folding involves arithmetic on group
elements of the underlying curve. For this to be efficient inside the
circuit, the group's base field must equal the circuit's native field.
Since the circuit's native field is the scalar field of the curve on which
the prover commits the SNARK --- which in turn is generally different from the
curve's own base field --- a difficulty arises: an $E_p$-IPA proof's group
elements have coordinates in $\F_p$, so verifying it natively would require
a circuit whose native field is $\F_p$. But a SNARK's verifier circuit
naturally operates over the curve's \emph{scalar} field $\F_q$ (where
$q = |E_p|$): the Fiat--Shamir challenges and folding scalars are $\F_q$
elements. Curves with $p = q$ do exist --- the \emph{anomalous} curves,
with trace of Frobenius $1$ --- but on them the discrete logarithm is
computable in polynomial time (Smart; Semaev; Satoh--Araki, 1998--99), so
no \emph{secure} curve has $\F_p = \F_q$, and no one curve $E_p$ can host
both the proof's coordinates and its own verifier circuit natively.

The resolution is a \emph{2-cycle} of elliptic curves, where $|E_p| = q$ and
$|E_q| = p$. Then:
\begin{itemize}[leftmargin=*]
\item A circuit committed on $E_q$ verifies a proof generated over $E_p$;
that circuit's native field is the scalar field of $E_q$, namely $\F_p$
(since $|E_q| = p$) --- which coincides with the coordinate field of
$E_p$, enabling native manipulation of $E_p$'s group elements.
\item A circuit over $\F_q$ verifies the next proof, generated over $E_q$.
\item Proofs alternate between the two curves.
\end{itemize}
This is precisely the Pasta cycle of the \emph{Math Guide},
where the two curves are Pallas (over $\F_{p_{\Pallas}}$, order $p_{\Vesta}$)
and Vesta (over $\F_{p_{\Vesta}}$, order $p_{\Pallas}$).

\section{The Halo 2 proof system}\label{sec:halo2}

We now assemble the abstractions of the \emph{Crypto Guide}--\S\ref{sec:pcd} into
the concrete proof system Halo 2: a transparent PLONKish SNARK whose
polynomial commitment is the inner-product argument over the Pasta cycle.
This section develops Halo 2 to the level of detail required to state
\emph{what an Orchard Action proof actually is}, the subject of
the \emph{Orchard Guide}.

\subsection{PLONKish arithmetisation}\label{sec:plonkish}

\begin{definition}[PLONKish constraint system]\label{def:plonkish}
Fix a prime field $\F$ and an integer $k$ defining a multiplicative
subgroup $H := \{\omega^0, \dots, \omega^{n-1}\} \subset \F^*$ of size
$n := 2^k$, where $\omega$ is a primitive $n$-th root of unity. A
\emph{PLONKish circuit} consists of:
\begin{itemize}[leftmargin=*]
\item Columns partitioned into \emph{fixed} (preprocessed),
\emph{advice} (prover-witness), and \emph{instance} (public-input).
\item A maximum constraint degree $d$.
\item A finite sequence of \emph{custom gates}
$g_t(\vec x_0, \vec x_1, \dots) = 0$ --- multivariate polynomials of
degree $\le d$ in column values at the current row and at offsets
$\omega^r X$ for fixed small $r$.
\item A subset $S \subseteq H \times [\#\text{cols}]$ of
\emph{equality-constraint cells} controlled by a permutation $\sigma$.
\item A set of \emph{lookup arguments}: input tuples $(A_0, \dots, A_{m-1})$
and table tuples $(S_0, \dots, S_{m-1})$, with the assertion that for
every row $i \in [0, n)$, $(A_0(\omega^i), \dots, A_{m-1}(\omega^i))$
matches some row of the table.
\end{itemize}
\end{definition}

Vanilla PLONK (Gabizon--Williamson--Ciobotaru, 2019) employs a fixed five-selector
universal gate
\[
q_M(X) a(X) b(X) + q_L(X) a(X) + q_R(X) b(X) + q_O(X) c(X) + q_C(X) \equiv 0 \pmod{Z_H(X)},
\]
where $Z_H(X) := X^n - 1$ is the \emph{vanishing polynomial} of the domain
$H$ (it is zero exactly on $H$, so ``$\equiv 0 \bmod Z_H$'' means ``zero at
every row''). ``PLONKish'' generalises this by permitting designer-chosen
gates of higher arity and degree, each gated by a selector $q_i(X)$
vanishing on rows where the gate is inactive. The Orchard Action circuit
uses ten advice columns plus a small fixed-column lookup table for
Sinsemilla; its custom gates encode the algebraic constraints of the
Action statement (the \emph{Orchard Guide}) together with the
complete- and incomplete-addition arithmetic for in-circuit
elliptic-curve operations (the \emph{Math Guide} names the distinction).

\paragraph{Polynomial representation.}
The Lagrange basis $\ell_j(X)$ for $H$, where $\ell_j(\omega^j) = 1$
and $\ell_j(\omega^{j'}) = 0$ for $j' \ne j$, interpolates each column into
a polynomial of degree $< n$ over $H$. The prover commits to each
polynomial via the IPA-PCS (\S\ref{sec:ipa-detail}). All subsequent
arguments are \emph{polynomial identities} on these committed polynomials,
verifiable by querying them at random points.

\subsection{From statement to circuit: arithmetisation in practice}
\label{sec:statement-to-circuit}

Definition~\ref{def:plonkish} states what a PLONKish circuit \emph{is}; it
does not say how a statement \emph{becomes} one. This subsection traces that
passage through the deployed code: a real custom gate (the ECC chip's
incomplete point addition), two real lookup declarations (the ten-bit range
check and the Sinsemilla generator table), and the synthesis machinery that
turns witness data into the column vectors on which every argument of this
section operates. Our sources are the crates the Orchard circuit is built
from --- \texttt{halo2\_proofs}~0.3.2 and \texttt{halo2\_gadgets}~0.4.0, as
resolved by \texttt{orchard}~0.13.1, the version we cite --- with file
paths relative to each crate's \texttt{src/}.

\paragraph{Shape first, data second.}
A circuit, to \texttt{halo2\_proofs}, is a type implementing the
\texttt{Circuit} trait, whose two principal methods split the construction
into phases (\texttt{halo2\_proofs}, \texttt{plonk/circuit.rs}). The static
method \texttt{configure} receives a mutable \texttt{ConstraintSystem} and
declares the \emph{shape}: it mints fixed, advice, and instance columns,
allocates selectors, registers gates and lookups, and marks columns as
equality-enabled. No witness exists at this point; \texttt{configure} is not
even handed an instance of the circuit type. The method \texttt{synthesize}
runs later, with the data, and fills cells. The split is the proof system's
division of labour: everything \texttt{configure} declares is public and
determines the verifying key --- the verifier's entire view of the circuit
--- while everything \texttt{synthesize} writes into advice cells is the
prover's witness. (Key generation also runs \texttt{synthesize} once,
against a witness-free backend that records fixed-cell values, selector
activations, and copy constraints while ignoring advice; the prover runs the
same code against a backend that records only advice. One synthesis
procedure, two observers: \texttt{Assembly} in \texttt{plonk/keygen.rs} and
\texttt{WitnessCollection} in \texttt{plonk/prover.rs}.)

\paragraph{Gates are queried identities.}
A call \texttt{meta.create\_gate(name, closure)} defines one custom gate.
Inside the closure, each \texttt{query\_advice(column, Rotation(r))} --- and
its fixed and instance analogues --- yields a symbolic handle to ``the value
of this column $r$ rows below the gate's own row'': the offsets $\omega^r X$
of Definition~\ref{def:plonkish}, realised at evaluation time by querying
the column polynomial at $\omega^r x$ (\texttt{halo2\_proofs},
\texttt{poly/domain.rs}, \texttt{rotate\_omega}). The closure combines the
handles by field arithmetic into an expression tree
(\texttt{Expression}: constants, selector nodes, queried cells, negations,
sums, products, scalings), one tree per constraint, and the helper
\texttt{Constraints::with\_selector} multiplies every constraint $c$ by the
gate's selector expression, storing $q \cdot c$
(\texttt{plonk/circuit.rs}). This helper is the mechanical origin of the
``selector times polynomial'' shape of every deployed gate but one --- the
generalisation, gate by gate, of the fixed vanilla-PLONK equation of
\S\ref{sec:plonkish}. (The exception, the ECC chip's witness-point gate,
multiplies by hand: its $(q \cdot x) \cdot (y^2 - x^3 - 5)$ and the
helper's $q \cdot \bigl( x \cdot (y^2 - x^3 - 5) \bigr)$ are equal
polynomials but different expression trees, and the pinned verifying key
freezes the tree; \texttt{halo2\_gadgets},
\texttt{ecc/chip/witness\_point.rs}.) A selector is virtual at this
stage: a named boolean
column, off everywhere by default, enabled row by row during synthesis. At
key generation, \texttt{compress\_selectors} packs the selectors into
genuine fixed columns --- several to a column wherever the circuit's degree
bound leaves headroom --- and rewrites every selector node into a
fixed-column expression (\texttt{plonk/circuit.rs},
\texttt{plonk/keygen.rs}). The deployed Action circuit declares $56$
selectors, whose compressed columns are among the $29$ fixed columns pinned
in its verifying key; the pinned count is the post-compression one
(\texttt{orchard}~0.13.1, \texttt{src/circuit\_description}; since 0.14
the file lives at \texttt{src/circuit\_data/\allowbreak
circuit\_description\_fixed}).

\paragraph{Regions, the layouter, and wiring.}
Data enters at synthesis through a \texttt{Layouter}. A chip requests a
\emph{region} --- a small rectangular window of cells addressed by relative
offsets --- fills it cell by cell, and leaves the choice of the region's
absolute starting row to the \emph{floor planner}
(\texttt{halo2\_proofs}, \texttt{circuit.rs}). Values flow between regions
not by position but by \emph{copy constraints}: the call
\texttt{enable\_equality(column)} enrols a column in the single global
permutation of \S\ref{sec:perm}, and \texttt{constrain\_equal} --- or the
assign-and-constrain combination \texttt{copy\_advice} --- merges the two
cells' cycles in that permutation
(\texttt{plonk/\allowbreak permutation/\allowbreak keygen.rs}; the halo2 book's
\emph{Permutation argument} chapter documents the cycle-merging). Public
inputs bind the same way, an advice cell copy-constrained to an instance
cell (\S\ref{sec:instance}). The deployed circuit uses the measuring
\texttt{V1} floor planner, which records every region's shape in a
witness-free pass, sorts the regions by advice area, and first-fits the
largest ones first (\texttt{circuit/\allowbreak
floor\_planner/\allowbreak v1.rs}; the sort and first-fit in
\texttt{v1/\allowbreak strategy.rs}); the
\texttt{orchard} crate additionally pins the planner's sort to a vendored
Rust~1.56.1 \texttt{pdqsort} (feature
\texttt{floor-planner-v1-legacy-pdqsort}) so that the layout --- and with it
the consensus verifying key --- is identical across platforms and
toolchains.

\paragraph{A deployed gate: incomplete point addition.}
The ECC chip's incomplete-addition gate (\texttt{halo2\_gadgets},
\texttt{ecc/chip/add\_incomplete.rs}) computes $P + Q = R$ on Pallas under
the promises $P, Q \neq \mathcal{O}$ and $x_p \neq x_q$. Its configuration
is one selector $q_{\text{add-incomplete}}$ and four equality-enabled advice
columns \texttt{x\_p}, \texttt{y\_p}, \texttt{x\_qr}, \texttt{y\_qr}; the
gate queries $x_p, y_p, x_q, y_q$ at the current row and $x_r, y_r$ at the
next row of the \emph{same} two columns \texttt{x\_qr}, \texttt{y\_qr}
(hence their names: $Q$ on the selector row, $R$ below it). The two
constraints, named \texttt{x\_r} and \texttt{y\_r} in the source, are
\begin{align}
q_{\text{add-incomplete}} \cdot
  \bigl( (x_r + x_q + x_p)\,(x_p - x_q)^2 - (y_p - y_q)^2 \bigr) &= 0,
  \label{eq:add-inc-xr}\\
q_{\text{add-incomplete}} \cdot
  \bigl( (y_r + y_q)\,(x_p - x_q) - (y_p - y_q)\,(x_q - x_r) \bigr) &= 0.
  \label{eq:add-inc-yr}
\end{align}
These are the affine chord formulas (the \emph{Math Guide}) with the slope
eliminated: for $x_p \neq x_q$ they are equivalent to
$x_r + x_q + x_p = \lambda^2$ and $y_r + y_q = \lambda\,(x_q - x_r)$ with
$\lambda := (y_p - y_q)/(x_p - x_q)$, multiplied through by
$(x_p - x_q)^2$ and $(x_p - x_q)$ respectively to clear the denominators
(the halo2 book's \emph{Addition} chapter records the same clearing). The
raw degrees are three and two; with the selector factor, four and three ---
well inside the deployed circuit's constraint-degree bound of nine (below).
The \emph{complete}-addition companion gate, which branches over the
identity, doubling, and inverse cases that this gate excludes, needs twelve
constraint polynomials of degree up to six and five auxiliary witness cells
per invocation (\texttt{ecc/chip/add.rs}); incompleteness is a bargain
wherever its promises can be kept.

Synthesis for one invocation (\texttt{assign\_region}, same file) makes the
witness path concrete. The chip enables the selector at the region's row
$i$; copies the four input coordinates into that row with
\texttt{copy\_advice}, which both assigns each cell and copy-constrains it
back to its source; computes, \emph{out of circuit},
\[
\lambda = \frac{y_q - y_p}{x_q - x_p}, \qquad
x_r = \lambda^2 - x_p - x_q, \qquad
y_r = \lambda\,(x_p - x_r) - y_p;
\]
and assigns $x_r, y_r$ into row $i+1$ (Table~\ref{tab:add-inc-region}). The
slope $\lambda$ is never written to any cell: the prover uses it to compute
the answer, and the committed
identities~\eqref{eq:add-inc-xr}--\eqref{eq:add-inc-yr} verify the answer
without it. The division is not even performed here --- coordinates are
carried as deferred fractions (\texttt{halo2\_proofs},
\texttt{plonk/assigned.rs}), and every denominator falls to a single batch
inversion when synthesis ends (\texttt{plonk/prover.rs},
\texttt{batch\_invert\_assigned}). The output cells are returned as a new
non-identity point with no further on-curve check --- sound because, once
the call site guarantees distinct $x$-coordinates on non-identity inputs,
the two identities force $R = P + Q$, a non-identity curve point.

\begin{table}[htbp]
\centering
\begin{tabular}{l cccc c}
\toprule
row & \texttt{x\_p} & \texttt{y\_p} & \texttt{x\_qr} & \texttt{y\_qr} &
  $q_{\text{add-incomplete}}$ \\
\midrule
$i$     & $x_p$ & $y_p$ & $x_q$ & $y_q$ & $1$ \\
$i+1$   & --    & --    & $x_r$ & $y_r$ & $0$ \\
\bottomrule
\end{tabular}
\caption{The incomplete-addition region: two rows by four advice columns
(\texttt{halo2\_gadgets}, \texttt{ecc/chip/add\_incomplete.rs}). The four
input cells $x_p, y_p, x_q, y_q$ on the selector row $i$ are
copy-constrained to the cells they came from; the result cells $x_r, y_r$
occupy the \texttt{x\_qr}, \texttt{y\_qr} columns of row $i+1$, where the
gate's \texttt{Rotation::next()} queries find them. Dashes mark cells the
gate does not use.}
\label{tab:add-inc-region}
\end{table}

\paragraph{Incompleteness, and where it is discharged.}
Reading the two polynomials at the excluded inputs shows exactly what
``incomplete'' costs. If $Q = -P \neq \mathcal{O}$ (so $x_p = x_q$ and
$y_p = -y_q \neq 0$), constraint~\eqref{eq:add-inc-xr} collapses to
$-(2 y_p)^2 \neq 0$: no assignment satisfies the gate --- a completeness
failure, never a soundness one. If $Q = P$, both polynomials vanish
identically and $(x_r, y_r)$ is \emph{unconstrained}: a reachable doubling
would let the prover claim any result. If an input were the identity in the
gadget's $(0,0)$ affine encoding, the constraints would accept a garbage
output. The deployed circuit discharges the three exclusions in three
different places. Non-identity travels with the type: the gate's inputs
are \texttt{NonIdentityEccPoint} values. A fresh point entering the chip
is witnessed under the gate
\[
q_{\text{point-non-id}} \cdot \bigl( y^2 - x^3 - 5 \bigr) = 0
\qquad
(\texttt{halo2\_gadgets}, \texttt{ecc/chip/witness\_point.rs}),
\]
which $(0,0)$ cannot satisfy because $5$ is not a square and $-5$ is not a
cube in $\F_{p_{\Pallas}}$ (the halo2 book's \emph{Addition} chapter) ---
but no fresh witness ever reaches incomplete addition. In the deployed
circuit the incomplete-addition gate fires only inside fixed-base scalar
multiplication (\texttt{ecc/chip/mul\_fixed.rs}), which sums precomputed
windowed multiples $[(k+2) \cdot 8^w]B$ of a fixed base $B$. There the
addends are assembled by the type's unchecked constructor
(\texttt{from\_coordinates\_unchecked}) and pinned instead to the
precomputed window table by \texttt{mul\_fixed}'s own coordinate gates
(\texttt{coords\_check}, same file): the $x$-coordinate must equal a
degree-seven polynomial in the window value whose fixed coefficients
interpolate the window's eight table $x$-coordinates, and the
$y$-coordinate must satisfy $y + z_w = u^2$ against a per-window fixed
$z_w$ and a witnessed $u$, alongside an on-curve check --- together
pinning the point to the table entry selected by the window digit, and no
entry of the table is the identity. The running accumulator is a sum of
such entries under this very gate; the type carries the non-identity
guarantee from either construction site to the gate, which re-checks
nothing. The distinctness $x_p \neq x_q$ is enforced by no constraint at
all; it too is structural: the window tables store $[(k+2) \cdot 8^w]B$
rather than the naive $[(k+1) \cdot 8^w]B$ precisely because the naive
offset admits a window collision that produces the fatal doubling (the
halo2 book's \emph{Fixed-base scalar multiplication} chapter).
Prover-side, \texttt{assign\_region} rejects a \emph{known} bad witness
with an error --- a developer diagnostic, not a constraint. (The
circuit's other incomplete additions --- the variable-base
ladder's fused double-and-add rounds and Sinsemilla's two per chunk, the
ones the \emph{Orchard Guide}'s gadget survey counts --- are separate gates
built on the same discipline; \texttt{ecc/chip/mul/\allowbreak incomplete.rs},
\texttt{sinsemilla/chip.rs}.)

\paragraph{Non-algebraic relations become lookups.}
A gate can assert only polynomial identities. ``This value has ten bits''
and ``this point is the $j$-th Sinsemilla generator'' are not identities but
set memberships, and the circuit declares them as lookups: a call
\texttt{meta.lookup(closure)} returns pairs of an input \emph{expression}
and a table column, and pushes one lookup argument onto the constraint
system (\texttt{halo2\_proofs}, \texttt{plonk/circuit.rs}). Each
declaration is one instance of the argument of \S\ref{sec:lookup}, the
input expressions playing $A_0, \dots, A_{m-1}$ and the table columns
$S_0, \dots, S_{m-1}$. The deployed circuit carries exactly three
(\texttt{orchard}~0.13.1, \texttt{src/circuit\_description}): one shared
range-check declaration, and the Sinsemilla generator declaration twice ---
once per Sinsemilla chip instance, each on its own half of the ten advice
columns.

All three consume the same table. Loaded once at synthesis, it holds, for
$0 \le j < 2^{10}$,
\[
\text{row } j:\quad \bigl( j,\; X(S[j]),\; Y(S[j]) \bigr),
\qquad
S[j] :=
\GroupHash\bigl(\texttt{"z.cash:SinsemillaS"},\, \mathrm{LE}_{32}(j)\bigr),
\]
in three table columns $(t_{\mathrm{idx}}, t_x, t_y)$; the generators ship
as precomputed constants that the crate's own tests re-derive from the
hash (\texttt{sinsemilla} crate, \texttt{src/sinsemilla\_s.rs};
\texttt{halo2\_gadgets}, \texttt{utilities/\allowbreak
lookup\_range\_check.rs} for the
loading). After the assignment, the layouter fills every remaining usable
row of each table column with that column's row-$0$ value
(\texttt{halo2\_proofs}, \texttt{circuit/\allowbreak
table\_layouter.rs}); a defaulted
input row therefore still matches a genuine table row, and the type
\texttt{TableColumn} hides its underlying fixed column exactly so that no
chip can bypass this default-filling.

The range check is one declaration serving two modes
(\texttt{halo2\_gadgets},
\texttt{utilities/\allowbreak lookup\_range\_check.rs}):
\begin{equation}
\label{eq:lookup-range}
q_{\mathrm{lookup}} \cdot \Bigl(
  q_{\mathrm{running}} \cdot \bigl( z_i - 2^{10}\, z_{i+1} \bigr)
  + (1 - q_{\mathrm{running}}) \cdot z_i
\Bigr)
\;\longmapsto\; t_{\mathrm{idx}},
\end{equation}
with $z_i$ and $z_{i+1}$ the running-sum advice column at the current and
next rotations. With $q_{\mathrm{running}} = 1$ the input is a running-sum
step: the prover witnesses $z_{i+1} = (z_i - a_i) \cdot 2^{-10}$ on
successive rows, the expression recovers the limb
$a_i = z_i - 2^{10} z_{i+1}$, and membership in $t_{\mathrm{idx}} =
\{0, \dots, 2^{10}-1\}$ asserts that each limb of the decomposition has ten
bits; a \emph{strict} decomposition additionally copy-constrains the final
$z$ to the constant $0$, range-constraining the decomposed element to
$10w$ bits for $w$ limbs. With $q_{\mathrm{running}} = 0$ the cell itself
is looked up (checks of fewer than ten bits add a small bit-shift gate in
the same file). Every range check in the Action circuit funnels through
this one declaration, on a single advice column.

The Sinsemilla declaration is the three-column instance
(\texttt{halo2\_gadgets},
\texttt{sinsemilla/\allowbreak chip/\allowbreak generator\_table.rs}):
\begin{equation}
\label{eq:lookup-sinsemilla}
\Bigl(\;
q_{S1} \cdot m,\;\;
q_{S1} \cdot x_p + (1 - q_{S1}) \cdot X(S[0]),\;\;
q_{S1} \cdot y_p + (1 - q_{S1}) \cdot Y(S[0])
\;\Bigr)
\;\longmapsto\;
(t_{\mathrm{idx}},\, t_x,\, t_y),
\end{equation}
where $m = z_i - 2^{10} \cdot q_{\mathrm{run}} \cdot z_{i+1}$ recovers the
ten-bit message word from the hash's own running-sum decomposition (with
$q_{\mathrm{run}} := q_{S2} - q_{S2}(q_{S2} - 1)$, an expression in a fixed
column $q_{S2} \in \{0, 1, 2\}$ that evaluates to $1$ on the interior rows
of a message piece and to $0$ on final rows, where the trailing running sum
is implicitly zero), and where $y_p$ is not a witnessed cell at all but the
derived expression $Y_A \cdot 2^{-1} - \lambda_1 (x_A - x_p)$ over the
chip's double-and-add columns (\texttt{ecc/chip/mul/\allowbreak
incomplete.rs} supplies
$Y_A := (\lambda_1 + \lambda_2)(x_A - x_R)$ and
$x_R := \lambda_1^2 - x_A - x_p$). Lookup inputs, in other words, may be
arbitrary expressions over queried cells, not merely cells; the chip
proves that the generator it is adding \emph{is} $S[m]$ without ever
spending an advice cell on $y_{S[m]}$. The $(1 - q_{S1})$ branches default
every row on which the chip is inactive to table row $0$, so the membership
assertion of \S\ref{sec:lookup} holds on all rows. One API restriction is
visible here: the selectors gating lookup inputs
($q_{\mathrm{lookup}}$, $q_{\mathrm{running}}$, $q_{S1}$) must be declared
as \texttt{complex\_selector}s --- simple selectors are reserved for gates,
where the key-generation compression is licensed to rewrite them
(\texttt{halo2\_proofs}, \texttt{plonk/circuit.rs}).

\paragraph{The handoff.}
When \texttt{synthesize} returns, the construction has produced nothing
exotic: every column --- advice, fixed, instance --- is a vector of
$n = 2^k$ field elements; in the advice columns the deferred fractions are
resolved by one batch inversion and the reserved trailing rows filled with
blinding randomness (\S\ref{sec:halo-zk}, Remark~\ref{rem:active-rows}).
The prover commits to
each advice column directly in the Lagrange basis
(\texttt{halo2\_proofs}, \texttt{poly/commitment.rs},
\texttt{commit\_lagrange} --- the same group element as committing the
interpolated coefficients, so the commitment even precedes interpolation)
and then interpolates each column over $H$ by an inverse FFT
(\texttt{poly/domain.rs}, \texttt{lagrange\_to\_coeff}): these are
precisely the column polynomials of \S\ref{sec:plonkish}. Each gate's
stored expression tree is then evaluated with the committed column
polynomials substituted for the queried cells --- a query at rotation $r$
reading the polynomial at $\omega^r X$ --- and the results are folded with
powers of the challenge $y$ and divided by $Z_H(X) = X^n - 1$
(\texttt{plonk/prover.rs}; \texttt{plonk/vanishing/prover.rs};
\texttt{divide\_by\_vanishing\_poly} in \texttt{poly/domain.rs}). The gate
identities declared in \texttt{configure} have become the quotient $h(X)$
of Phase~4 (\S\ref{sec:halo2-full}). This is the point where the toy of
\emph{How Halo 2 Proves} resumes --- its \S\,\emph{From the grid to
polynomials} interpolates a four-row trace and assembles exactly this
$G(X)$ whose divisibility by $Z_H$ the rest of the protocol certifies ---
and where the \emph{Orchard Guide}'s gadget stack bottoms out: each gadget
of its survey is a bundle of \texttt{configure}-time gates and lookups plus
\texttt{synthesize}-time regions, and after synthesis nothing remains of
the stack but rows in these same columns. The scale-up from the toy is
numerical, not conceptual. The deployed Action circuit has $k = 11$
($n = 2048$ rows), the ten advice columns of \S\ref{sec:plonkish}, one
instance column, and twenty-nine fixed columns, with a constraint-degree
bound of nine across
its gates and permutation and lookup expressions --- so the quotient splits
into $9 - 1 = 8$ chunks, and the prover's FFTs run over an extended domain
of size $2^{14} = 8 \cdot 2^{11}$ (\texttt{orchard}~0.13.1,
\texttt{src/circuit\_description}; \texttt{halo2\_proofs},
\texttt{poly/domain.rs}). The next two subsections supply the arguments
this construction leans on: the permutation argument behind every copy
constraint, and the lookup argument behind
declarations~\eqref{eq:lookup-range}
and~\eqref{eq:lookup-sinsemilla}.


\subsection{The permutation (equality) argument}\label{sec:perm}

A PLONK-style permutation argument, which Halo 2 generalises to multiple
columns, enforces equality constraints between distinct cells. Each
cell $(i, j)$ (column $i$, row $j$) receives a unique label
$\delta_i \cdot \omega^j$ with $\delta_i$ chosen so all products are
distinct as $(i, j)$ ranges over the trace. The prover writes
$\sigma$-permuted polynomials $s_i(X)$ with $s_i(\omega^j) =
\delta_{i'} \cdot \omega^{j'}$ when $\sigma(i, j) = (i', j')$.

Given Fiat--Shamir challenges $\beta, \gamma \in \F$, the prover commits
to a running-product polynomial $Z_{\mathrm{perm}}(X)$ asserting, where
$v_i(X)$ denotes the Lagrange interpolation (\S\ref{sec:plonkish}) of the
$i$-th column participating in the permutation,
\[
\prod_{i, j} \frac{v_i(\omega^j) + \beta\,\delta_i\,\omega^j + \gamma}
{v_i(\omega^j) + \beta\,s_i(\omega^j) + \gamma} = 1.
\]
The assignment satisfies every equality constraint (cells related by
$\sigma$ carry equal values) precisely when the denominator factors are a
permutation of the numerator factors, in which case the product equals $1$
for every $\beta, \gamma$; conversely, if some constraint is violated, the
product at the sampled $\beta, \gamma$ equals $1$ with probability at most
$O(N/|\F|)$ by Schwartz--Zippel, where $N$ is the number of cells
participating in the permutation. The verifier checks this by opening $Z_{\mathrm{perm}}$
at $x$ and $\omega x$, verifying the boundary $Z_{\mathrm{perm}}(\omega^0) = 1$
and the row-by-row update. The per-proof verifier work is $O(m + \log n)$,
where $m$ denotes the number of columns participating in the permutation.

\subsection{The lookup argument}\label{sec:lookup}

Halo 2's lookup argument is structurally simpler than plookup
(Gabizon--Williamson 2020) and supports arbitrary (possibly non-fixed)
tables.

\paragraph{Objective.} Establish that every entry of column $A(X)$ on the active
domain appears as some entry of column $S(X)$.

\paragraph{Construction.}
The prover supplies polynomials $A'(X), S'(X)$ where (i) $A'$ is a
permutation of $A$ with like values grouped into contiguous runs, and
(ii) $S'$ is a permutation of $S$ in which the first row of each run of
like values in $A'$ has the matching value in $S'$. The prover commits to
$A'$ and $S'$; after receiving challenges $\beta, \gamma$, it commits to
a running-product column $Z_{\mathrm{lookup}}(X)$ and proves:
\begin{align*}
\ell_0(X) \bigl(1 - Z_{\mathrm{lookup}}(X)\bigr) &= 0,\\
Z_{\mathrm{lookup}}(\omega X)(A'(X) + \beta)(S'(X) + \gamma) - Z_{\mathrm{lookup}}(X)(A(X) + \beta)(S(X) + \gamma) &= 0,\\
\ell_0(X)(A'(X) - S'(X)) &= 0,\\
(A'(X) - S'(X))(A'(X) - A'(\omega^{-1} X)) &= 0.
\end{align*}
The first two constraints, with the boundary conditions of
Remark~\ref{rem:active-rows}, force $A'$ and $S'$ to be permutations of
$A$ and $S$ respectively; the third aligns row $0$ ($A'_0 = S'_0$), where
the wrap-around reference $A'(\omega^{-1}X)$ deprives the fourth
constraint of force. The fourth constrains every subsequent row either to
start a new run aligned with $S'$ ($A'_i = S'_i$) or to continue the
previous one ($A'_i = A'_{i-1}$). Inductively, every value of $A$ equals
some value of $S'$, hence of $S$.

A random-linear combination with a challenge $\theta$ reduces multi-column
lookups to single-column:
$A_{\mathrm{c}}(X) := \sum_j \theta^{m-1-j} A_j(X)$. Variable tables
(where $S$ is itself an advice column) and range checks
(where $S$ enumerates the allowed range) are special cases.

\begin{remark}[Active rows and blinding]\label{rem:active-rows}
The lookup constraints above, like the permutation constraints of
\S\ref{sec:perm}, hold on the \emph{active} rows only: the deployed
system reserves the trailing rows of every column for random blinding
values (\S\ref{sec:halo-zk}). Each product-update constraint is therefore
multiplied by the factor
$\bigl(1 - (q_{\mathrm{last}}(X) + q_{\mathrm{blind}}(X))\bigr)$, where
the fixed columns $q_{\mathrm{blind}}$ and $q_{\mathrm{last}}$ select the
blinding rows and the boundary row immediately before them, and boundary
constraints such as
$\ell_{\mathrm{last}}(X)\bigl(Z(X)^2 - Z(X)\bigr) = 0$ pin the running
product to $\{0, 1\}$ at the boundary row. The random rows consequently
neither violate the arguments nor weaken them.
\end{remark}

\subsection{The inner-product polynomial commitment}\label{sec:ipa-detail}

The Halo 2 polynomial commitment, instantiating the abstract PCS of
the \emph{Crypto Guide}, is the inner-product argument of
Bootle--Cerulli--Chaidos--Groth--Petit (EUROCRYPT 2016) as refined in
Bulletproofs (B\"unz et al., 2017). We give the construction in detail.
Write $d := 2^k$ for the degree bound (equivalently, the number of
generators); the round count $k = \log_2 d$ reuses the letter $k$ of
Definition~\ref{def:plonkish}, harmlessly, because in the assembled
protocol of \S\ref{sec:halo2-full} the committed polynomials have degree
$< n$, so $d = n = 2^k$ there. The letter $H$ likewise does double duty,
inherited from two literatures: $H \subset \F$ is the PLONK evaluation
domain (as in $Z_H$), while $H \in \G$ is the Pedersen blinding generator
of the \emph{Crypto Guide}; the ambient set ($\F$ versus $\G$)
disambiguates every occurrence. Two further letters carry double duty:
$p$ denotes the group order (as in $\F_p$) and, with an argument or in
$\deg p$, the committed polynomial $p(X)$; and the permutation
polynomials $s_i(X) \in \F[X]$ of \S\ref{sec:perm} are distinct from the
IPA structured scalars $s_i \in \F_p$ of \S\ref{sec:gpoly}. Type and
argument disambiguate every occurrence.

\begin{construction}[IPA-based polynomial commitment]\label{const:ipa}
Let $\G$ be a cyclic group of prime order $p$ in which DL is hard. Choose
$d = 2^k$ and sample, via hash-to-curve, $\vec{G} = (G_1, \dots, G_d)$
and $H \in \G$ with no known non-trivial linear relation among them.

For $p(X) = \sum_{i=0}^{d-1} a_i X^i \in \F_p[X]$ and blinder
$r \in \F_p$,
\[
\Comm(p; r) := \ip{\vec{a}}{\vec{G}} + [r] H \in \G.
\]
\end{construction}

To open $p$ at $x \in \F_p$ with claimed value $v$, the relation is
\[
\Rel_{\mathrm{IPA}} := \Bigl\{\bigl((P, x, v); (\vec{a}, r)\bigr) :
P = \ip{\vec{a}}{\vec{G}} + [r] H,\;
v = \ip{\vec{a}}{\vec{b}(x)},\; \deg p \le d-1\Bigr\},
\]
with $\vec{b}(x) := (1, x, x^2, \dots, x^{d-1})$.

\subsubsection{The recursive halving protocol}

The argument runs in $k = \log_2 d$ rounds. Initialise
$\vec{a}^{(k)} := \vec{a}$, $\vec{G}^{(k)} := \vec{G}$,
$\vec{b}^{(k)} := \vec{b}(x)$, and let $U \in \G$ be a challenge group
element (in Fiat--Shamir, derived from the transcript). The prover
absorbs $[v]U$ so that the relation becomes
\[
P + [v] U = \ip{\vec{a}}{\vec{G}} + [r] H + [\ip{\vec{a}}{\vec{b}}] U.
\]

In round $j$ (descending from $k$ to $1$): split each vector into halves
and let
\begin{align*}
L_j &:= \ip{\vec{a}_{\mathrm{lo}}^{(j)}}{\vec{G}_{\mathrm{hi}}^{(j)}} + [\ell_j] H + [\ip{\vec{a}_{\mathrm{lo}}^{(j)}}{\vec{b}_{\mathrm{hi}}^{(j)}}] U,\\
R_j &:= \ip{\vec{a}_{\mathrm{hi}}^{(j)}}{\vec{G}_{\mathrm{lo}}^{(j)}} + [r_j] H + [\ip{\vec{a}_{\mathrm{hi}}^{(j)}}{\vec{b}_{\mathrm{lo}}^{(j)}}] U.
\end{align*}
The verifier returns a uniform $u_j \in \F_p^*$. Both parties fold:
\begin{align*}
\vec{a}^{(j-1)} &:= u_j\,\vec{a}_{\mathrm{lo}}^{(j)} + u_j^{-1}\,\vec{a}_{\mathrm{hi}}^{(j)},\\
\vec{G}^{(j-1)} &:= u_j^{-1}\,\vec{G}_{\mathrm{lo}}^{(j)} + u_j\,\vec{G}_{\mathrm{hi}}^{(j)},\\
\vec{b}^{(j-1)} &:= u_j^{-1}\,\vec{b}_{\mathrm{lo}}^{(j)} + u_j\,\vec{b}_{\mathrm{hi}}^{(j)}.
\end{align*}

After $k$ rounds each vector has length 1. The prover sends the final
scalar $a := \vec{a}^{(0)}$ and an aggregated blinder $r^*$. The verifier
accepts iff
\begin{equation}\label{eq:ipa-final}
P + [v] U + \sum_{j=1}^{k} \bigl([u_j^2] L_j + [u_j^{-2}] R_j\bigr)
\;\stackrel{?}{=}\; [a]\,G^{(0)} + [r^*] H + [a \cdot b^{(0)}] U,
\end{equation}
where $G^{(0)} = \ip{\vec{s}}{\vec{G}}$ and the challenges determine
$\vec{s}$ as below.

\subsubsection{The structured scalars and the polynomial $g$}\label{sec:gpoly}

Write the binary expansion $i = \sum_{\ell=0}^{k-1} b_\ell 2^\ell$ for
$i \in [0, d)$. By induction on the folding,
\[
G^{(0)} = \ip{\vec{s}}{\vec{G}}, \qquad b^{(0)} = \ip{\vec{s}}{\vec{b}(x)} = g(x; u_1, \dots, u_k),
\]
where
\[
s_i := \prod_{\ell=0}^{k-1} u_{\ell+1}^{(-1)^{1-b_\ell}}, \qquad
g(X; u_1, \dots, u_k) := \prod_{\ell=1}^{k} \bigl(u_\ell^{-1} + u_\ell X^{2^{\ell-1}}\bigr).
\]

The essential asymmetry is the following: $g(X)$ has degree $d - 1$ and the verifier can evaluate
it at $x$ in $O(\log d)$ field operations; but computing
$G^{(0)} = \Comm(g; 0)$ requires an $O(d)$-size multi-scalar
multiplication (MSM). This asymmetry is the seed of the accumulation
scheme (\S\ref{sec:halo-acc}).

\subsubsection{Worked example: an IPA opening at \texorpdfstring{$d = 4$}{d=4}}\label{sec:ipa-worked}

Take $d = 4$, so $k = 2$ rounds. Let $p(X) = a_0 + a_1 X + a_2 X^2 + a_3 X^3$
with $\vec{a} = (a_0, a_1, a_2, a_3)$. The verifier holds
$P = \ip{\vec{a}}{\vec{G}} + [r] H$ (we suppress $r$ to declutter; the
$H$-part propagates identically) and requires $v = p(x)$ at some $x \in \F_p$.

Set $\vec{b}^{(2)} := (1, x, x^2, x^3)$ and $\vec{G}^{(2)} := \vec{G}$.
After the prover absorbs $[v] U$, the relation is
$P + [v] U = \ip{\vec{a}}{\vec{G}} + \ip{\vec{a}}{\vec{b}^{(2)}}\,U$.

\paragraph{Round 2.} Split into halves of size 2. The prover sends
\begin{align*}
L_2 &= a_0 G_3 + a_1 G_4 + (a_0 x^2 + a_1 x^3) U,\\
R_2 &= a_2 G_1 + a_3 G_2 + (a_2 + a_3 x) U.
\end{align*}
The verifier picks $u_2 \in \F_p^*$ and both parties fold to length-2
vectors as above.

\paragraph{Round 1.} Split the length-2 vectors. The prover sends $L_1,
R_1$ analogously; the verifier picks $u_1$. After folding, all three vectors
have length 1.

\paragraph{Final check.} With $i = b_0 + 2 b_1$ ranging over
$\{0, 1, 2, 3\}$:
\begin{align*}
s_0 &= u_1^{-1} u_2^{-1}, & s_1 &= u_1 u_2^{-1}, & s_2 &= u_1^{-1} u_2, & s_3 &= u_1 u_2.
\end{align*}
The verifier evaluates
$g(x; u_1, u_2) = (u_1^{-1} + u_1 x)(u_2^{-1} + u_2 x^2)$ in four
field multiplications ($u_1 x$, $x^2$, $u_2 x^2$, and the product of the
two factors), and computes
$G^{(0)} = \sum_{i=0}^{3} [s_i] G_{i+1}$ as a length-4 MSM. The check
(\ref{eq:ipa-final}) becomes
\[
P + [v] U + [u_1^2] L_1 + [u_1^{-2}] R_1 + [u_2^2] L_2 + [u_2^{-2}] R_2
\;\stackrel{?}{=}\; [a] G^{(0)} + [a \cdot g(x; u_1, u_2)]\,U.
\]

The folding rule for $\vec{G}$ ensures that $G^{(0)}$ is a specific
linear combination of the original $\vec{G}$ with coefficients $\vec{s}$;
that the $s_i$ match $\prod_\ell (u_\ell^{-1} \text{ or } u_\ell)$
according to the binary expansion of $i$ is exactly the closed-form
identity. The verifier can evaluate $g$ in $O(\log d)$ field operations
(about $3k-2$: $k$ coefficient products, $k-1$ squarings of $x$, $k-1$
factor multiplications) but recomputing $G^{(0)}$ takes $O(d)$ MSM ---
the asymmetry the accumulator defers.

\subsubsection{Knowledge soundness via rewinding}\label{sec:ipa-extract}

\begin{theorem}[Knowledge soundness of IPA]\label{thm:ipa-ks}
Under the DL assumption on $\G$, and modelling Fiat--Shamir as a random
oracle, Construction~\ref{const:ipa} is knowledge-sound for
$\Rel_{\mathrm{IPA}}$ with knowledge error $O(d / |\F_p|)$.
\end{theorem}

\begin{proof}[Sketch of the tree extractor]
The extractor rewinds the prover $\mathcal{P}^*$ to obtain, at each
round, \emph{three} accepting transcripts that agree on $(L_j, R_j)$ but
carry pairwise distinct challenges $u_j$. The round-$j$ verification
equation involves the three monomials $u_j^{-2}, 1, u_j^2$, whose
coefficients --- the unknown representations of $L_j$, the folded
commitment, and $R_j$ over the generators --- the extractor must
determine; two transcripts would leave this system underdetermined.
Three challenges with pairwise distinct squares (\(u \ne \pm u'\); the
colliding pairs are excluded below) yield three linear equations whose
\(3 \times 3\) Vandermonde-type determinant in \(u^2\) is non-zero, so
the extractor solves for the
round-$j$ witness halves $\vec{a}_{\mathrm{lo}}^{(j)},
\vec{a}_{\mathrm{hi}}^{(j)}$ (Lemma~\ref{lem:single-round}).

Extracting the full witness $\vec{a} \in \F_p^d$ requires three sibling
transcripts at the deepest level, which yield one round-1 witness; three
such triples, hung off pairwise distinct challenges at the next-shallower
round, yield one round-2 witness; and so on, each shallower round
tripling the requirement. The tree thus has
$3^k = d^{\log_2 3}$ leaves --- polynomial in $d$ (\S\ref{sec:ks}) ---
yielding polynomial expected extraction time when $d = \poly(\lambda)$.
The theorem's knowledge error of $O(d / |\F_p|)$ does not come from a
tree-wide union bound. Each of the $k = \log_2 d$ rounds excludes $O(1)$
bad challenges ($u_j = \pm u_j'$), contributing
$O(\log d / |\F_p|) \le O(d / |\F_p|)$; Schwartz--Zippel terms of degree
$d$ account for the rest. Collisions among the resampled challenges
inside the extractor's $3^k$-leaf tree affect only its expected running
time --- the extractor rewinds and resamples --- not the knowledge error;
a naive union bound over the roughly $3^k$ sibling challenge draws would
bound the tree-wide collision probability by
$O(d^{\log_2 3} / |\F_p|)$, still negligible, but that is not the source
of the theorem's figure.

This is the recursive generalisation of special soundness for Schnorr
(the \emph{Crypto Guide}): a single transcript reveals nothing, but a few
transcripts on the same first message permit inversion of a small linear
system; in the IPA the system is $3 \times 3$ per round and the recursion
stacks $k$ rounds.
\end{proof}
\subsection{Accumulation and the Halo trick}\label{sec:halo-acc}

The verifier's check (\ref{eq:ipa-final}) runs in $O(\log d)$ scalar
operations and $\log d$ point additions, \emph{except} for the single
evaluation $G^{(0)} = \ip{\vec{s}}{\vec{G}}$ --- an MSM of length $d$.
The accumulation idea (Bowe--Grigg--Hopwood 2019; formalised by BCMS20)
is to \emph{defer} this MSM and amortise it across many proofs.

\begin{definition}[Atomic accumulator for IPA]\label{def:ipa-acc}
An IPA \emph{accumulator instance} is a tuple
$\Acc := (G^{(0)}, u_1, \dots, u_k) \in \G \times \F_p^k$. It is
\emph{valid} iff
$G^{(0)} = \ip{\vec{s}(\vec{u})}{\vec{G}} = \Comm(g(X; u_1, \dots, u_k);\, 0)$
--- the deferred claim concerns the unblinded commitment, the blinder
having been discharged by the $[r^*] H$ term of (\ref{eq:ipa-final}).
Subsequent unary $\Comm(\cdot)$ abbreviates $\Comm(\cdot\,; 0)$.
\end{definition}

\begin{theorem}[Accumulation of IPA, BCMS20]\label{thm:bcms20-acc}
There is an accumulation scheme
$(\mathsf{AccVerifier}, \mathsf{AccDecider})$ for IPA with:
\begin{itemize}[leftmargin=*]
\item $\mathsf{AccVerifier}$, on input a new IPA instance and an old
accumulator, runs in $O(\log d)$ and outputs a new accumulator $\Acc'$.
\item $\mathsf{AccDecider}$, run \emph{once} at the end of a chain, performs
an MSM of length $d$ to check $G^{(0)} = \Comm(g)$.
\item Soundness reduces to DL on $\G$ in the random-oracle model.
\end{itemize}
\end{theorem}

\paragraph{Operational interpretation.}
A Halo 2 verifier circuit, expressed over the scalar field of one Pasta
curve and verifying a proof produced over the \emph{other} curve, can:
\begin{enumerate}[leftmargin=*]
\item Run the $O(\log d)$ ``succinct'' checks of the IPA verifier natively.
\item Witness the next-step claimed $G^{(0)\prime}$ and the next challenges
$u_1', \dots, u_k'$ as public inputs to the next proof, deferring the MSM
check.
\item Fold the deferred MSM checks: with a random challenge $\rho$, the
two claims $G^{(0)} = \Comm(g)$ and $G^{(0)\prime} = \Comm(g')$ reduce to
the single check $G^{(0)} + [\rho]\, G^{(0)\prime} = \Comm(g + \rho g')$,
with both challenge tuples retained so that the combined length-$d$
coefficient vector $\vec{s}(\vec{u}) + \rho\, \vec{s}(\vec{u}\,')$ is
recomputable; a false claim survives the combination with probability at
most $1/|\F_p|$. In the recursive instantiation the combined commitment
is opened at a fresh random point inside the next proof's multipoint
argument --- each $g_i$ there is evaluable in $O(\log d)$ --- so the
deferred claim emerging from the next IPA opening is again a single tuple
of the form of Definition~\ref{def:ipa-acc}; the tuple type is closed
under accumulation only through this prover-assisted step, not under
verifier-side linear combination.
\end{enumerate}
A single final MSM at the end of the chain validates the entire history.
The Pasta cycle (\S\ref{sec:cycle}) renders each step native: a
Vesta-IPA proof's circuit, whose native field is the Vesta scalar field
$\F_{p_{\Pallas}}$, verifies the previous Pallas-IPA proof, whose group
elements (the column and quotient commitments, the $(L_j, R_j)$ pairs,
the accumulator's $G^{(0)}$) have coordinates in exactly that field; the
next Pallas proof
verifies the Vesta proof by the symmetric argument. This cross-curve
alternation is what enables proof-carrying data without expensive
non-native field simulation. The same coincidence is what lets Orchard's
$\F_{p_{\Pallas}}$-native Action circuit manipulate the Pallas points
$\cmsym$, $\rk$, and $\cvsym^{\mathsf{net}}$ directly
(\S\ref{sec:instance}).

Orchard's deployment does not recurse: the verifier of an Action proof
pays the deferred MSM itself, as part of the standard verification cost.
Even there the accumulation idea is load-bearing: a Zcash node verifying
many Action proofs takes a random linear combination of their deferred
MSMs and evaluates the combination as one MSM, amortising the linear-time
work across the whole batch --- the deployed, non-recursive use of the
same algebraic structure.

The mathematical core is the recognition that
$G^{(0)} = \Comm(g(X; u_1, \dots, u_k))$ --- the linear-time check is
\emph{itself a polynomial commitment opening of a structured polynomial}
and hence amenable to recursion within an IPA.

\subsection{The complete Halo 2 protocol}\label{sec:halo2-full}

We now assemble the pieces into the seven-phase protocol implemented in
the halo2 codebase. Throughout, $\F$ denotes the constraint field
(Orchard: $\F_{p_{\Pallas}}$), $\G$ the IPA group (Orchard:
$\Vesta(\F_{p_{\Vesta}})$), and $H$ the multiplicative subgroup of $\F$
of order $n$.

\paragraph{Setup (preprocessed, one-time).}
The circuit designer specifies columns, custom gates, lookup tables, and
the equality permutation $\sigma$. Both parties derive commitments to the
fixed-column polynomials $f_i(X)$, selector polynomials $q_i(X)$,
permutation polynomials $s_i(X)$, and lookup-table polynomials $S_j(X)$,
together with the Lagrange basis on $H$, the vanishing polynomial
$Z_H(X) := X^n - 1$, and the IPA generators $\vec{G} \in \G^n$ and
$H \in \G$. The commitments to the fixed, selector, permutation, and lookup-table
polynomials, together with the domain parameters, constitute the
\emph{verifying key} $\mathsf{vk}$; every part of it is recomputable from
the public circuit description, which is why no trusted setup exists. To
verify a proof one needs exactly three inputs: $\mathsf{vk}$, the
instance (public-input) values, and the proof itself.

\paragraph{Phase 1 --- advice commitments.}
Both parties first absorb a digest of $\mathsf{vk}$ and the (unblinded)
instance commitments into the transcript (\S\ref{sec:instance}), so every
subsequent challenge depends on the circuit and the public input. The
prover then constructs the advice column values $\vec{w_i} \in \F^n$
satisfying all gates and equality constraints; fills the reserved
trailing rows with uniformly random field elements (the blinding rows of
\S\ref{sec:halo-zk}); interpolates each $\vec{w_i}$ to a polynomial
$w_i(X)$ of degree $< n$ on $H$; commits via IPA-PCS with a fresh
blinder; sends the commitments.

\paragraph{Phase 2 --- lookup compression: challenge $\theta$.}
The verifier (via Fiat--Shamir, here and throughout) returns
$\theta \in \F$. For each lookup the prover compresses multi-column
inputs and tables with powers of $\theta$, computes the permuted columns
$A', S'$ (\S\ref{sec:lookup}), and commits to them.

\paragraph{Phase 3 --- running products: challenges $\beta, \gamma$.}
The verifier returns $\beta, \gamma \in \F$. The prover commits to the
permutation running product $Z_{\mathrm{perm}}$ (\S\ref{sec:perm}) and to
one lookup running product $Z_{\mathrm{lookup}}$ per lookup
(\S\ref{sec:lookup}). The order matters for soundness: the permuted
columns $A', S'$ must be fixed \emph{before} the challenges
$\beta, \gamma$ that randomise the products are drawn.

\paragraph{Phase 4 --- quotient challenge $y$.}
The verifier returns a fresh $y \in \F$. The prover assembles the
\emph{gate equation}: a single polynomial $C(X)$ that is a $y$-weighted
linear combination of (i) each custom gate, (ii) permutation constraints,
(iii) lookup constraints. By construction $C(X)$ vanishes on $H$ when
all gates pass, so the quotient
$h(X) := C(X) / Z_H(X) = C(X) / (X^n - 1)$ is a polynomial. Its degree,
however, is bounded by \(d(n-1) - n = (d-1)(n-1) - 1\) for maximum gate
degree \(d\) --- a gate of degree \(d\) multiplies up to \(d\) column
polynomials of degree \(< n\), and division by \(Z_H\) removes \(n\) ---
which exceeds the IPA's degree bound $n$. The prover therefore
splits $h$ into \(\lceil (\deg h + 1)/n \rceil\) chunks
$h_0, h_1, \dots$ of degree $< n$, with
$h(X) = \sum_i X^{n i}\, h_i(X)$, and commits to each chunk separately;
the verifier recombines the chunk evaluations at the evaluation point.
The prover's FFTs correspondingly run over an \emph{extended} domain
whose size accommodates $\deg C$, a small multiple of $n$.

\paragraph{Phase 5 --- evaluation challenge $x$.}
The verifier returns $x \in \F$. The prover evaluates every committed
polynomial at $x$, and at the rotated points the constraints reference
--- $\omega x$ for the running products (and for any gate reading an
adjacent row), $\omega^{-1} x$ for the permuted lookup input $A'$ ---
and sends these claimed evaluations as field elements. The verifier checks
the gate equation $C(x) = h(x) Z_H(x)$ as a numerical identity in the
sent evaluations. This check is meaningful only once the claimed
evaluations bind to the commitments, which is the role of phases 6--7;
a prover who sent inconsistent evaluations would pass this field identity
but fail the opening argument.

\paragraph{Phase 6 --- multipoint opening.}
The claimed evaluations must now be bound to the commitments, and Halo 2
reduces all the required openings to a single IPA opening. The committed
polynomials are grouped by their set of query points: most are queried at
$\{x\}$ alone, while the running products are also queried at $\omega x$
(and $A'$ at $\omega^{-1} x$), so only a handful of distinct point sets
arise. Within each group, a challenge $x_1$ folds the polynomials into a
single combination $q_i(X)$ by powers of $x_1$. (The letters $q$ and $f$
here follow the halo2 book's multipoint-opening notation and are
unrelated to the selector and fixed-column polynomials of the Setup
paragraph, which are themselves among the polynomials being folded; the
index $i$ now ranges over query-point sets, not columns.) For each group, both
parties interpolate the low-degree polynomial $r_i(X)$ that agrees with
the claimed evaluations on the group's point set; the quotient
$f_i(X) := \bigl(q_i(X) - r_i(X)\bigr) / \prod_{z}(X - z)$, with $z$
ranging over the group's points, is a polynomial precisely when the
claimed evaluations are correct. The prover combines the $f_i$ by powers
of a second challenge $x_2$ into $f(X)$ and commits to it. The verifier
returns a fresh evaluation point $x_3$, and the prover sends the
evaluation of each $q_i$ at $x_3$; the verifier reconstructs the expected
value of $f(x_3)$ from the $q_i(x_3)$, the $r_i(x_3)$, and the vanishing
factors --- if each $f_i$ is a genuine polynomial this value matches the
committed $f$, and by Schwartz--Zippel a match at random $x_3$ convinces
the verifier that every claimed evaluation was correct. A final challenge
$x_4$ folds $f$ and the $q_i$ into one polynomial $f^*(X)$, whose
expected evaluation at $x_3$ the verifier assembles from the same data;
a single opening of $f^*$ at $x_3$ then discharges every original
opening at once.

\paragraph{Phase 7 --- IPA opening.}
Prover and verifier execute the recursive halving IPA
(\S\ref{sec:ipa-detail}) on $f^*$, yielding $\log_2 n$ pairs $(L_j, R_j)$,
a final scalar $a$, and a final blinder $r^*$. The verifier accepts iff
(\ref{eq:ipa-final}) holds.

\paragraph{Accumulation (recursion).}
The next proof's verifier circuit \emph{defers} the MSM check
$G^{(0)} = \Comm(g(X; u_1, \dots, u_k))$ at the end of
phase 7 into an accumulator and folds it with prior
accumulators (\S\ref{sec:halo-acc}).
A single decider at the very end validates the entire chain.

\paragraph{The concrete transcript.}
The deployed implementation instantiates Fiat--Shamir with a transcript
built on the BLAKE2b hash function. Every prover message ---
commitments as curve points, claimed evaluations as field elements --- is
absorbed into the hash state in the order the protocol fixes, and each
verifier challenge is squeezed from the state at its prescribed point,
after everything it must depend on has been absorbed. Both parties also
absorb the data they share in advance --- the $\mathsf{vk}$ digest and
the unblinded instance commitments --- which therefore never appear in
the proof. The proof \emph{is} the ordered sequence of the prover's
messages; the verifier replays the absorptions and recomputes every
challenge.

\paragraph{Costs.}
The prover's work is dominated by FFTs over the extended domain and by
one $O(n)$-size MSM \emph{per committed polynomial} --- several tens in
all: each advice column, the permuted lookup columns, the running
products, the quotient chunks, and the multipoint polynomial $f$
(typical Orchard Action circuit: $n = 2^{11}$, a few seconds on commodity
hardware). The verifier performs $O(\log n)$ field and group operations
per proof --- recomputing the challenges, the gate-equation field check,
the $\log n$ rounds of phase 7 --- plus the single deferred $O(n)$ MSM,
amortised over a batch as in \S\ref{sec:halo-acc}.

\paragraph{Proof size.}
Two kinds of data make up a proof. The IPA of phase 7 contributes
$2 \log_2 n$ group elements (the pairs $(L_j, R_j)$) plus the final
scalars --- for Orchard's $n = 2^{11}$, twenty-two points; this is the
only part that scales with $n$, and only logarithmically. Everything
else scales with the \emph{circuit structure}, independently of $n$: one
commitment per advice column, two permuted-column commitments and one
running product per lookup, the permutation-product and quotient-chunk
commitments, the multipoint commitment, and one field element per claimed
evaluation. Each Pasta point or field element serialises to 32 bytes.
For the Orchard Action circuit the total is fixed by consensus (ZIP 225):
a bundle's proof occupies $2720 + 2272 \cdot a$ bytes for $a$ Actions ---
just under 5\,kB for a single-Action bundle --- the per-Action term
covering the per-circuit commitments and evaluations, the constant term
the shared quotient, multipoint, and IPA data.


\subsection{Instance columns and binding the public statement}\label{sec:instance}

Definition~\ref{def:plonkish} lists \emph{instance} columns alongside fixed
and advice columns, but the arguments of \S\ref{sec:perm}--\S\ref{sec:halo2-full}
never single them out. We single them out now, because what an Orchard Action proof
\emph{asserts} is precisely a statement about its instance values --- the
anchor, nullifier $\nfsym$, net value commitment $\cvsym^{\mathsf{net}}$,
randomised key $\rk$, extracted new-note commitment $\cmx$, and the
$\mathsf{enableSpends}/\mathsf{enableOutputs}$
flags (the \emph{Orchard Guide}).

\paragraph{The instance polynomial.}
An instance column is a public-input column: its $n$ entries
$\vec{\pi} = (\pi_0, \dots, \pi_{n-1}) \in \F^n$ are not part of the
witness but both parties agree on them before verification. Like every
other column it is interpolated over $H$ in the Lagrange basis,
\[
\pi(X) := \sum_{j=0}^{n-1} \pi_j\, \ell_j(X) \in \F[X], \qquad \deg \pi < n,
\]
where $\ell_j(\omega^j) = 1$ and $\ell_j(\omega^{j'}) = 0$ otherwise
(\S\ref{sec:plonkish}). The prover writes the public values into a handful of
designated rows; all remaining $\pi_j$ are $0$.

\paragraph{Binding by copy constraints.}
Orchard binds the public values through the permutation argument of
\S\ref{sec:perm}: the instance column participates in the permutation
$\sigma$, and the circuit equality-constrains each instance cell to the
advice cell carrying the corresponding in-circuit value (halo2's
\texttt{constrain\_instance}). A satisfying assignment therefore exists
\emph{only} when the advice cells holding the anchor, $\nfsym$,
$\cvsym^{\mathsf{net}}$, $\rk$, $\cmx$, and the flags equal the public values the
verifier supplied. (An alternative design binds public inputs with a
dedicated gate $q(X)\bigl(a(X) - \pi(X)\bigr) = 0$, where the selector
$q$ is active exactly on the designated rows; Orchard does not use it ---
the copy constraint reuses the existing permutation machinery at no extra
gate degree.)

\paragraph{Where it enters the protocol.}
The verifier never takes $\pi(X)$ on trust from the prover. In the
deployed IPA backend, \emph{both} parties compute the commitment to each
instance polynomial --- unblinded, since the values are public, so the
two computations agree --- and absorb it into the transcript before the
first challenge (Phase~1, \S\ref{sec:halo2-full}); every Fiat--Shamir
challenge thereby depends on the public input. Thereafter the instance
column is treated like any other: its claimed evaluation $\pi(x)$ appears
among the Phase~5 evaluations, enters the gate equation and the
permutation expressions, and the multipoint opening of Phase~6 binds it
to the instance commitment. Because the verifier computed that commitment
itself, from the instance values it intends to verify against, the prover
has no freedom over $\pi$: an evaluation inconsistent with the verifier's
own values fails the opening argument.

\paragraph{Consequence.}
A prover who targets different public values $\vec{\pi}\,'$ must open
against the instance commitment the \emph{verifier} computes from the
genuine $\vec{\pi}$, and must satisfy the copy constraints against those
values. Knowledge soundness
(\S\ref{sec:ks}) then extracts a witness for the relation instantiated at
exactly those public values; a prover that does not know such a witness
--- whatever witnesses it holds for other instances --- convinces the
verifier with at most negligible probability. This binds the proof to its public
statement: an Action proof that a node verifies against an anchor it accepts,
a nullifier it will mark spent, and a value commitment it will sum into
the net balance check (the \emph{Orchard Guide}) can only verify if the
prover knew a witness consistent with \emph{those} values.

\subsection{Zero knowledge: hiding the witness in Halo 2}\label{sec:halo-zk}

Knowledge soundness (\S\ref{sec:ks}) establishes that a convincing prover \emph{knows} a
witness; it remains to explain why the proof \emph{reveals} no more than that. Halo~2 is
zero-knowledge through two devices --- hiding commitments and blinded polynomials
--- assembled into a simulator.

\paragraph{Hiding commitments.} The polynomial commitment of
\S\ref{sec:ipa-detail} is \emph{perfectly hiding}. The prover commits a polynomial \(p\)
as
\[
  \mathrm{Commit}(p; r) \;=\; \ip{\vec{p}}{\vec{G}} + [r]\,H, \qquad
  r \stackrel{\$}{\leftarrow} \F_p,
\]
with \(\vec p\) the coefficient vector of \(p\) and \(H\) an independent generator
of no known discrete-log relation to \(\vec G\). The \([r]H\) term makes the
commitment a \emph{uniform} group element for every \(p\), so a commitment by
itself leaks nothing about what the prover committed (the perfect hiding of a Pedersen
commitment, \emph{Crypto Guide}). Every commitment the prover sends in
phases 1--6 --- the advice columns, the permuted lookup columns $A'$, $S'$,
the permutation and lookup running products, the quotient chunks, and the
multipoint polynomial $f$ --- has this form. The verifying-key and instance
commitments are deliberately unblinded: they commit public data both parties
recompute (\S\ref{sec:halo2-full}, \S\ref{sec:instance}), so there is
nothing to hide.

\paragraph{Blinding the witness polynomials.} The prover eventually
\emph{opens} a commitment, and an opening reveals evaluations. The prover interpolates the advice columns
over the domain \(H\) to polynomials of degree \(< n\); to prevent those
evaluations from leaking the witness, the prover does not use all \(n\) rows for
circuit data. It reserves a small fixed number of rows at the bottom of the table
--- the \emph{blinding rows}, never covered by an active gate --- and fills them
with uniform random field elements, at least one more than the number of times the
verifier will open any single column. Its \(n\) evaluations pin down a degree-\((<n)\) polynomial,
so with enough independent random values planted on the
blinding rows the handful the verifier actually queries are jointly uniform and
independent of the circuit data. The prover blinds the permutation and lookup running products the same
way; Remark~\ref{rem:active-rows} describes how the constraint system
exempts the blinding rows, so the random values never violate an
argument. The quotient pieces admit no such treatment --- $C$ fully
determines them --- so their commitment blinders, together with the
blinding-row randomness they inherit through $C$, protect them instead;
their single collapsed opening at $x$ is the publicly computable value
$C(x)/(x^n - 1)$. (The deployed implementation also commits a uniformly
random polynomial alongside the quotient and opens it at the same point,
keeping the multipoint argument zero-knowledge.)

\paragraph{Why the openings leak nothing.} In each phase the verifier sees only
(i) hiding commitments --- uniform group elements --- and (ii) a bounded number of
evaluations of the committed polynomials at the Fiat--Shamir challenge points. The
blinding rows make each such evaluation uniformly random subject only to the
\emph{public} relations the verifier itself checks; no information about the
private witness passes through. The final inner-product opening
(\S\ref{sec:ipa-detail}) likewise runs with a blinder, so it too hides its
input.

\paragraph{The simulator.} Formally, zero knowledge is the existence of a
\emph{simulator} that, given only the public input and \emph{not} the witness,
outputs a transcript indistinguishable from a real one. Halo~2's simulator runs
the protocol ``backwards,'' exactly as Schnorr's does (\emph{Crypto Guide}): it
chooses the evaluations and the final opening so as to satisfy the verifier's
equations, then chooses hiding commitments consistent with them --- possible
precisely because the commitments are perfectly hiding and it may choose the blinders
freely --- and, non-interactively, \emph{programs} the Fiat--Shamir oracle so
the challenges come out as required. Since the commitments are perfectly hiding
and a blinder protects every opening, the simulated transcript is distributed as a real
one. Interactively this is honest-verifier zero knowledge; compiled through
Fiat--Shamir in the random-oracle model it is a NIZK that reveals nothing beyond
the truth of the statement.

\paragraph{The three properties together.} For the relation fixed by a PLONKish
circuit, Halo~2 thus delivers all three guarantees a SNARK should:
\emph{completeness} (an honest prover holding a satisfying witness always
convinces the verifier), \emph{knowledge soundness} (\S\ref{sec:ks}: from any
convincing prover an extractor pulls a satisfying witness, so an accepted proof
certifies one is \emph{known}), and \emph{zero knowledge} (a witness-free
simulator reproduces the proof, so it leaks nothing else). It is exactly this
combination --- ``the prover knows a satisfying witness, and the proof discloses
nothing more'' --- on which the Orchard Action proof rests: knowledge soundness
yields balance and nullifier integrity, zero knowledge yields privacy
(\emph{Orchard Guide}).
\end{document}
